Algebraic File Synchronization: Adequacy and Completeness

### 2018-07-08 p2/axiom-proof.tex
Spellcheck

\begin{proof}
Suppose $\cxynnv$ and $\czwnv$ satisfy the conditions,
and there is an $\FS$ for which $(\cxynnv\cc\czwnv)\aFS\neq\fsbroken$.
If $\czwnv\neq\cbba{n}$, then either $\FS(n)\neq\empty$ or $\czwnv\aFS(n)\neq\empty$,
and therefore $\FS(\parentf{n})$ must be a directory for the filesystem to satisfy
the tree property at all times.
However, from $\cxynnv\neq\cdda{\nn}$ we also know that either $\FS(\nn)$ 
(before applying $\cxynv$) or $\cxynv\aFS(\nn)$ (after applying $\cxynv$)
is not a directory. Depending on which is true, we have a point in the sequence
of applying the commands where the filesystem cannot satify the tree property
and therefore becomes broken as $\parentf{n}$, a descendant of $\nn$, contains
a directory and so must all its ancestors.
\end{proof}

>>>>

\begin{proof}
Suppose $\cxynnv$ and $\czwnv$ satisfy the conditions,
and there is an $\FS$ for which $(\cxynnv\cc\czwnv)\aFS\neq\fsbroken$.
If $\czwnv\neq\cbba{n}$, then either $\FS(n)\neq\empty$ or $\czwnv\aFS(n)\neq\empty$,
and therefore $\FS(\parentf{n})$ must be a directory for the filesystem to satisfy
the tree property at all times.
However, from $\cxynnv\neq\cdda{\nn}$ we also know that either $\FS(\nn)$ 
(before applying $\cxynv$) or $\cxynv\aFS(\nn)$ (after applying $\cxynv$)
is not a directory. Depending on which is true, we have a point in the sequence
of applying the commands where the filesystem cannot satisfy the tree property
and therefore becomes broken as $\parentf{n}$, a descendant of $\nn$, contains
a directory and so must all its ancestors.
\end{proof}

### 2018-07-08 p2/axioms.tex
Spellcheck

\begin{mydef}[$\eqext$, $\strext$ and $\equiv$]
% Relation in the algebra
We write $A\eqext B$, or say that $B$ \emph{extends} $A$,
% intended interpretation:
to mean that they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS\in\Dom{A}:\,A\aFS=B\aFS$.
% old inference rule: $ A\eqext B \Rightarrow S\cc A\cc T\eqext S\cc B\cc T$,
We can also see that $A\eqext B$ and $S\eqext T$ implies $A\cc S\eqext B\cc T$.
% symmetric:     aRb <=> bRa
% antisymmetric: aRb & bRa => a=b
% reflexive:     aRa
% transitive:    aRb & bRC => aRc
% preorder:      reflexive, transitive
% partial order: reflexive, transitive, antisymmetric
% equivalence:   reflexive, transitive, symmetric
% congruence:    https://en.wikipedia.org/wiki/Congruence_relation
% precongruence: congruence of a preorder
In other words, $\eqext$ is a preorder, and also a precongruence.

@@

\myskip
Command pairs in genral have the form
\[ \cxynv\cc  \czwmv \]
where $x,y,z,w\in\setv$ and $n,m\in\setn$. 
Extending \cref{def:command_categories}, we call certain pairs
\emph{construction} or \emph{destruction pairs}.

@@

We can see that construction pairs consist of construction comands,
while destruction pairs consist of destruction commands.

>>>>

\begin{mydef}[$\eqext$, $\strext$ and $\equiv$]
% Relation in the algebra
We write $A\eqext B$, or say that $B$ \emph{extends} $A$,
% intended interpretation:
to mean that they behave in the same way
on any filesystem $A$ does not break:
$\forall \FS\in\Dom{A}:\,A\aFS=B\aFS$.
% old inference rule: $ A\eqext B \Rightarrow S\cc A\cc T\eqext S\cc B\cc T$,
We can also see that $A\eqext B$ and $S\eqext T$ implies $A\cc S\eqext B\cc T$.
% symmetric:     a R b <=> b R a
% antisymmetric: a R b & b R a => a = b
% reflexive:     a R a
% transitive:    a R b & b R C => a R c
% preorder:      reflexive, transitive
% partial order: reflexive, transitive, antisymmetric
% equivalence:   reflexive, transitive, symmetric
% congruence:    https://en.wikipedia.org/wiki/Congruence_relation
% precongruence: congruence of a preorder
In other words, $\eqext$ is a preorder, and also a precongruence.

@@

\myskip
Command pairs in general have the form
\[ \cxynv\cc  \czwmv \]
where $x,y,z,w\in\setv$ and $n,m\in\setn$. 
Extending \cref{def:command_categories}, we call certain pairs
\emph{construction} or \emph{destruction pairs}.

@@

We can see that construction pairs consist of construction commands,
while destruction pairs consist of destruction commands.

### 2018-07-08 p2/def-filesystem.tex
Spellcheck

The combination of the set of nodes and the parent function,
$\langle \setn, \parent \rangle$, forms a \emph{skeleton}
which the filesytems populate with values from $\setv$.
As we require filesystem functions to be total,
we use a special value, $\empt\in\setv$, to indicate that the filesystem
is empty at a particular node, that is, there are no files or directories there.
We also assume that a filesystem has finitely many non-empty nodes.

@@

An essential assumption in our model is that there is only
a \emph{single directory value} ($|\setd|=1$).
The realisation that
this creates a symmetry between directories and empty nodes
and makes it possible to exploit a hidden symmetry in filesystem commands
that allows formulating a correct and complete reconciliation algorithm
is one of the main contributions of this paper.
In \cref{sec_multidir} we explore the applicability of our model given this assumption:
why applicability may not be affected and ways to relax this restriction.

>>>>

The combination of the set of nodes and the parent function,
$\langle \setn, \parent \rangle$, forms a \emph{skeleton}
which the filesystems populate with values from $\setv$.
As we require filesystem functions to be total,
we use a special value, $\empt\in\setv$, to indicate that the filesystem
is empty at a particular node, that is, there are no files or directories there.
We also assume that a filesystem has finitely many non-empty nodes.

@@

An essential assumption in our model is that there is only
a \emph{single directory value} ($|\setd|=1$).
The realization that
this creates a symmetry between directories and empty nodes
and makes it possible to exploit a hidden symmetry in filesystem commands
that allows formulating a correct and complete reconciliation algorithm
is one of the main contributions of this paper.
In \cref{sec_multidir} we explore the applicability of our model given this assumption:
why applicability may not be affected and ways to relax this restriction.

### 2018-07-08 p2/multidir.tex
Spellcheck

Despite the above, there can be applications where directory metadata
is considered important.
Synchronizers based on our model can be readily extended to handle them
by duplicating $\setn$. 
Given the input filesystems over $\setn$, we add a special node as an extra child under each original node,
and we encode directory metadata in a file under each directory.
It is easy to see that update detection and conflict resolution can continue as expected,
with the only exception of a potential conflict detected on one of these special nodes.
In these cases the synchronizer may prescribe creating a directory without creating
the special metadata file, which is clearly not possible to do on the target filesystem,
as creating a directory entails specifying its metadata as well.
In these cases, the synchronizer can either fall back to default values and flag the issue
for review later, or if, as suggested above, conflicts on the metadata (e.g. readable, writeable and executable flags)
can be easily automated, then it could form part of the implementation.

Reintroducing the \emph{move} command can happen in a similar fashion.
Its main advantages include that 
\emph{move} commands are easier to review,
and if the synchronizer sugegsts or performs a move instead
of deleting a file and recreating it somewhere else, the user
can be assured that no information is lost.
During pre-processing, if an update detector is used,
\emph{move} commands still do not need to be part of the command sequences
the synchronizer operates on.
If the command sequences are derived from filesystem journals, they
can be broken into two updates to allow the existing reconciliation
algorithm to operate.

>>>>

Despite the above, there can be applications where directory metadata
is considered important.
Synchronizers based on our model can be readily extended to handle them
by duplicating $\setn$. 
Given the input filesystems over $\setn$, we add a special node as an extra child under each original node,
and we encode directory metadata in a file under each directory.
It is easy to see that update detection and conflict resolution can continue as expected,
with the only exception of a potential conflict detected on one of these special nodes.
In these cases the synchronizer may prescribe creating a directory without creating
the special metadata file, which is clearly not possible to do on the target filesystem,
as creating a directory entails specifying its metadata as well.
In these cases, the synchronizer can either fall back to default values and flag the issue
for review later, or if, as suggested above, conflicts on the metadata (e.g. readable, writable and executable flags)
can be easily automated, then it could form part of the implementation.

Reintroducing the \emph{move} command can happen in a similar fashion.
Its main advantages include that 
\emph{move} commands are easier to review,
and if the synchronizer suggests or performs a move instead
of deleting a file and recreating it somewhere else, the user
can be assured that no information is lost.
During pre-processing, if an update detector is used,
\emph{move} commands still do not need to be part of the command sequences
the synchronizer operates on.
If the command sequences are derived from filesystem journals, they
can be broken into two updates to allow the existing reconciliation
algorithm to operate.

### 2018-07-08 p2/ordering.tex
Resolve TODOs
Spellcheck

% TODO Commuting commands important as we need to establish order

@@

As we recognize that mulitple orders may be possible,
let us describe our ordering algorithm by
defining a subset of the permutations, $\ordersetsign$.
The simplicity of the definition means that enumerating
all sequences in $\ordersetsign$ can be implemented by a straightforward
algorithm.

@@

\begin{proof}
% BACK_TO_THE_FS TODO
Formally, we prove our proposition by contradiction,
and in this version of the proof we reach back to our filesystem model.
We assume $\wrks{S}$ and therefore there is a $\FS$ for which $S\aFS\neq\fsbroken$,
but $S$ contains commands on a node and its parent that do not form a construction or destruction pair.
Select two such commands, $\cxynv$ and $\czwmv$,
and split $S$ around them into three parts:
\[ S = S_0 \cc \cxynv \cc S_1 \cc \czwmv \cc S_2. \]
As $S$ is simple, there cannot be any commands on $n$ or $m$ in $S_1$,
and therefore $(S_0\cc\cxynv\cc S_1)\FS(n) = y$,
and $[(S_0\cc\cxynv\cc S_1)\FS(m)] = [S_0\FS(m)]$.
If $[z]\neq[S_0\FS(m)]$, then $S\aFS=\fsbroken$, which is a contradiction.

@@

\begin{mylem}\mlabel{equiv_simple_same_commands}
If two simple sequences $A$ and $B$ are equivalent
and do not break all filesystems ($A\equiv \wrksx{B}$),
then they must contain the same commands.
\end{mylem}
\begin{proof}
% BACK_TO_THE_FS
% TODO Rule?
Let $A$ and $B$ be simple sequences
such that $A\equiv \wrksx{B}$,
and $\FS$ be a filesystem that they are defined on.
We use a proof by contradiction, and assume that they do not contain the same commands.
Without loss of generality, we can assume
that $A$
contains $\cxynv$, and $B$ either contains a different command
$\czwnv$ on $n$, or no command on $n$ at all.
As $A$ is simple, we know that $\cxynv$ is not an assertion command,
and therefore either $[x]\neq [y]$, or it is a replacement command.

@@

% TODO delete unused results

>>>>

@@

As we recognize that multiple orders may be possible,
let us describe our ordering algorithm by
defining a subset of the permutations, $\ordersetsign$.
The simplicity of the definition means that enumerating
all sequences in $\ordersetsign$ can be implemented by a straightforward
algorithm.

@@

\begin{proof}
% BACK_TO_FS
Formally, we prove our proposition by contradiction,
and in this version of the proof we reach back to our filesystem model.
We assume $\wrks{S}$ and therefore there is a $\FS$ for which $S\aFS\neq\fsbroken$,
but $S$ contains commands on a node and its parent that do not form a construction or destruction pair.
Select two such commands, $\cxynv$ and $\czwmv$,
and split $S$ around them into three parts:
\[ S = S_0 \cc \cxynv \cc S_1 \cc \czwmv \cc S_2. \]
As $S$ is simple, there cannot be any commands on $n$ or $m$ in $S_1$,
and therefore $(S_0\cc\cxynv\cc S_1)\FS(n) = y$,
and $[(S_0\cc\cxynv\cc S_1)\FS(m)] = [S_0\FS(m)]$.
If $[z]\neq[S_0\FS(m)]$, then $S\aFS=\fsbroken$, which is a contradiction.

@@

\begin{mylem}\mlabel{equiv_simple_same_commands}
If two simple sequences $A$ and $B$ are equivalent
and do not break all filesystems ($A\equiv \wrksx{B}$),
then they must contain the same commands.
\end{mylem}
\begin{proof}
% BACK_TO_FS
Let $A$ and $B$ be simple sequences
such that $A\equiv \wrksx{B}$,
and $\FS$ be a filesystem that they are defined on.
We use a proof by contradiction, and assume that they do not contain the same commands.
Without loss of generality, we can assume
that $A$
contains $\cxynv$, and $B$ either contains a different command
$\czwnv$ on $n$, or no command on $n$ at all.
As $A$ is simple, we know that $\cxynv$ is not an assertion command,
and therefore either $[x]\neq [y]$, or it is a replacement command.

@@
